The pragmatics of perspective-taking

### 2018-04-05 writing/journal_manuscript/manuscript.tex
Pass through manuscript; start on results

\begin{abstract}
%To communicate successfully, speakers must flexibly adapt %the specificity of 
%their utterances to the demands of the current context. When they are unsure exactly which objects are in view of their partner, for example, they may provide additional disambiguating information when referring to a particular object: saying "cassette tape" instead of the potentially ambiguous ``tape."
Recent debates over adult theory of mind use have hinged on behavior in natural-language communication tasks but have largely neglected the nuanced pragmatics of these tasks. In particular, using occluded objects to create perspective asymmetries may have unexpected side effects under pragmatic reasoning. Cooperative speakers ought to flexibly adjust their utterances to be more informative in contexts with high uncertainty about what is in the listener's field of view. In a direct replication, we show that scripted confederates violate these expectations by using less informative labels than unscripted directors naturally use. In a follow-up experiment, we directly manipulate whether objects are hidden to test the pragmatic account of speaker informativity in uncertain environments. Hence, apparent failures of visual perspective-taking may in fact be successes for sophisticated theory of mind use in context.
%Recent theories of language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication. To reconcile these points of view, we use a computational model of communication under uncertainty to analyze the pragmatic demands of a common theory of mind test.  Through a minimal pair of replications comparing scripted vs. unscripted speakers, we show that scripted confederates used in previous studies are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation, thus violating the listener's rational social expectations. This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind. 
\end{abstract}

@@

Our ability to successfully communicate and coordinate with social partners relies on some degree of contextual flexibility in production and interpretation. It may be technically correct to refer to a particular entity as a ``dog,'' but speakers tend to choose more specific or \emph{informative} expressions when trying to point out their dog at a busy dog park. Similarly, a hyperbolic statement like ``I waited a million years for a table'' tends to be interpreted as an expression of frustration at longer-than-expected wait times rather than a literal statement about the time elapsed \citep{KaoWuBergenGoodman14_NonliteralNumberWords}.  %A doctor who uses a single Latinate term to tell their colleague about a condition may use many more words to tell a patient the same thing.  % Our success as a cooperative species depends upon our ability to make sense of other agents. 
This systematic contextual variation could in principle be accommodated by internally caching a complex set of different meanings, but pragmatic accounts have argued for a more parsimonious explanation. Just as making sense of an agent's physical movements in the world requires an accurate model of how beliefs and intentions translate into motor plans \citep{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}, making sense of language depends on an accurate mental model of what a speaker would say, or what a listener would think, in different situations. In other words, flexible communication relies on a generic \emph{theory of mind} mechanism.

%While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, 
Past efforts to evaluate theory of mind in communication have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \citep{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}. 
We argue in this paper that the mental models supporting theory of mind use in communication are richer than previously explored. 
In particular, we consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \citep{FrankeJager16_ProbabilisticPragmatics, Grice75_LogicConversation,Clark96_UsingLanguage,GoodmanFrank16_RSATiCS, FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context. 
%For example, suppose a friend wants to refer to a nearby Dalmatian. 
%There are numerous descriptions that would all be literally true, but different expressions are preferred in different contexts.
%If there are no other dogs nearby, we may expect them to use a parsimonious basic-level term like ``dog.''
%When there is potential for confusion, however, we expect them to consider our likely interpretation and introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \citep{BrennanClark96_ConceptualPactsConversation, VanDeemter16_ComputationalModelsOfReferring, GrafEtAl16_BasicLevel}. 
%When speakers violate these expectations, listeners draw inferences about the intended meaning in order to account for the violation: for example, XXX \cite{XXX, BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}.

@@

This suggests that we consider whether the scripted utterances produced by  confederates in popular perspective-taking tasks were actually what a speaker in that context would be expected to say \citep{KuhlenBrennan13_LanguageInDialogue}. 
If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
More precisely, when both players know that objects are occluded in the display, speakers add additional precision to references in order to avoid confusion. 
If the listener justifiably expects the speaker to do so, they may \emph{rationally} adopt a strategy of choosing the \emph{a priori} more likely referent of the referring expression regardless of perspective. 
In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item. 
%\begin{table*}
%\begin{center}
%\begin{tabular}{ p{2cm} | r | r |  r || r | r | r || r | r | r}
%& \multicolumn{3}{c||}{\% attempted at least once} & \multicolumn{3}{c||}{\% attempted at least twice} & \multicolumn{3}{c}{\% of total cases}\\
%\hline
%& Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2  \\
%Experimental & 71 & 93 & 61 & 46 & 57 & 32 & 30 & 43 & 24\\
%Baseline        & 0   & 7   & 0   & 0   & 0   & 0   & 0   & 2   & 0\\
%\end{tabular}
%\caption{Side-by-side comparison of error rates in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications. The first two sections show the percentage of \emph{participants} attempting to move the occluded distractor at least once, or twice, of the four possible cases. The third section shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object. }
%\label{table:mainResults}
%\end{center}
%\end{table*}
%
%We began by replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment. We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study. We replicated the original finding, but noted a tendency of directors to be overinformative in unscripted filler trials. We then ran the same experiment without using any scripted instructions, observing unconstrained director utterances. We found much greater precision in unconstrained director utterances, which match targets much better than distractors, and better performance of the matchers. This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.

%In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners who are sensitive to the pragmatics of the situation expect these more informative utterances and produce ``errors'' when their expectations are flouted.
In Experiment 1, we conduct a direct replication of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, manipulating whether the speaker is a confederate or naive participant. We find that unscripted speakers naturally produce more informative utterances than those in the confederate's script. In Experiment 2, we test the hypothesis that speakers systematically produce longer, more specific utterances than strictly required by their visible context when they know objects are occluded. %Finally, in Experiment 3, we manipulate the listener's pragmatic beliefs about whether the speaker is indeed expected to produce such `over-informative' utterances. 
Taken together, these results suggest that apparent failures of perspective-taking are in fact attributable to sophisticated pragmatic expectations about speaker behavior---that is, to successful theory of mind use.

 \begin{figure}[t!]
 \begin{center}
 \vspace{-.25cm}
 \includegraphics[scale=.5]{figures/Exp1_design.png}
 \vspace{-.5cm}
 \caption{Screenshots of interface on a critical trial: a cassette tape is in view of both players, but a roll of tape is occluded from the speaker's view.}
 \vspace{-.5cm}
 \label{fig:exp1paradigm}
 \end{center}
 \end{figure}

@@

We recruited K participants (K/2 pairs) from Amazon Mechanical Turk to play an interactive communication game using the framework described in \cite{Hawkins15_RealTimeWebExperiments}. All participants were native English speakers. %Three pairs were excluded for making 2 or more errors on non-critical items.

@@

In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it. Furthermore, the matcher tended to spend more time hovering their cursor over the occluded cell, and less time hovering over the target object, in ambiguous experimental trials than in baseline trials with no ambiguity. Both of these results support Keysar's claim that adults do not reliably deploy theory of mind reasoning when it is called for.

%Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data. First,

@@

\section{Evaluating models of referring}

\subsection{Pure egocentrism}

The simplest strategy for a speaker attempting to refer to a target object $o_i$ is to produce the simplest label that best fits $o_i$, effectively ignoring context and listener knowledge. \todo[inline]{rxdh: technically, mean soft-max here...} Similarly, the simplest strategy for a listener is to pick the object in their view that is best described by the utterance, ignoring the visual access or pragmatic goals of the speaker. These models are purely egocentric, in the sense that they do not engage in any social reasoning.

\subsection{Pure consideration of visual access}

The most immediate model of social reasoning is consideration of the informativeness of a label in a shared context containing many objects. For a speaker, this means preferring labels that are good fits for the target but \emph{not} known distractors in common view. For instance, when referring to a dog in the shared context of other dogs, a speaker may opt to be more informative than simply saying `dog,' since it is ambiguous. For a listener, this means discounting objects that are not in shared context, regardless of how well the description fits. These are not available to the visual access of the speaker and thus not viable referents.

\subsection{Probabilistic weighing of perspectives}

Rather than taking a purely egocentric perspective or purely considering a partner's visual access, speakers or listeners may be probabilistically integrating these two perspectives \cite{HellerParisienStevenson16_ProbabilisticWeighing}. % When conversational expectations lead participants to expect over-informative utterances, they (rationally) place relatively less weight on features of the environment determining common ground, such as shared perceptual access.

\subsection{Gricean communication under uncertainty}

We begin by considering the problem faced by a speaker who wants their partner to select a target object from a context containing other potentially similar objects. %  a simple probabilistic model of reference in the rational speech act (RSA) model.

where speakers produce informative utterances by reasoning about a rational listener agent who updates their beliefs about the world according to the literal semantics of the language. This allows us to test pragmatic accounts of task performance.

Pragmatic critiques of the director-matcher task point to an additional source of uncertainty that has not previously been explored: uncertainty over a communication partner's visual context. Considering this uncertainty, of course, \emph{is} theory of mind in action; our goal is to show how such a model built on social reasoning can produce the `failures' found by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} as well as the phenomena observed by both \cite{HawkinsGoodman16_Keysar} and \cite{RubioFernandez16_DirectorTaskAttention}. For simplicity, we consider the full $2\times 2$ grid containing a single occluded cell, which was used by \cite{RubioFernandez16_DirectorTaskAttention}. We incorporate this source of uncertainty into both the speaker and listener models as follows:

\begin{enumerate}
\item A pragmatic speaker perceives an array of three objects $\mathcal{O}_S = \{o_i\}$, one of which is identified as a target $t$, but assumes that their partner sees an additional hidden object $o_h$ that is not in common ground. The speaker has uncertainty over the identity of this object, represented by a prior distribution $P(o_h)$ which they marginalize over when deciding what utterance to produce:
$$S_1(u | t, \mathcal{O}_p) \propto \sum_{o_h} L_0(t| \mathcal{O}_S \cup \{o_h\})P(o_h)P(u)$$

%\item A pragmatic listener, reasoning about this uncertain speaker, is presented with a full array of four objects $\mathcal{O}_L$, one of which is identified as hidden to the speaker, $o_h$. The listener, however, also has some uncertainty over whether the experimenter is deceptive, represented by a Bernoulli random variable $d$, such that they assign a small prior probability $P(d)$ to the possibility that $o_h$ is actually in common ground. Upon hearing an utterance$u$, they update both their beliefs about which object is the target and whether the experimenter is being deceptive:
%$$L_2(t, d | u, \mathcal{O}_L) = P(t)\left(S(u | t, \mathcal{O}_L)P(d)\right)^d(S(u | t, \mathcal{O}_L - o_h)(1-P(d))^{1-d}$$
\end{enumerate}
Technical details are omitted here for brevity, but the full model is fully specified in WebPPL and available to run at %\url{http://forestdb.org/models/keysar.html}.

%We focus on two critical patterns of results produced by the pragmatic listener $L_2$. First, we consider how it responds to the scripted instruction ``fish'' when presented with a context containing two fish (a red one that is occluded and a blue one in common ground) as well as two other unrelated objects. While it is most likely to choose the red fish in common ground, it has a relatively high probability (43\%) of choosing the hidden red fish, an effect \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} claimed was due to failure of theory of mind. This decision emerges in the model for two primary reasons. For one, a speaker with uncertainty over context could in principle be communicating about any object, and if they believed the hidden object was a red fish, the utterance `fish' would have a high probability of being produced. To a lesser extent, context uncertainty makes the overinformative utterance `blue fish' more likely than `fish', hence the omission of the color modifier has the (subtle) implication of \emph{not} the blue fish. This speaker phenomenon, while subtle under the parameter regime of the current implementation, produces the dynamic observed in the unscripted replication by  \cite{HawkinsGoodman16_Keysar}.
%
%Second, we consider the listener's posterior over the \emph{deception} Bernoulli variable $d$. Just as \cite{RubioFernandez16_DirectorTaskAttention} found that listeners were more likely to suspect deception after hearing an `overinformative' utterance like `blue fish' when only one fish was in common ground, so too does our pragmatic listener. The pragmatic speaker is relatively more likely to produce the additional color modifier if they see multiple fish than when they see a single fish and simply have uncertainty over the unknown object. Hence, our model simultaneously captures all three phenomena.

>>>>

\begin{abstract}
%To communicate successfully, speakers must flexibly adapt %the specificity of 
%their utterances to the demands of the current context. When they are unsure exactly which objects are in view of their partner, for example, they may provide additional disambiguating information when referring to a particular object: saying "cassette tape" instead of the potentially ambiguous ``tape."
Recent debates over adult theory of mind use have hinged on natural-language communication tasks with knowledge asymmetries, but have largely neglected the nuanced pragmatics of these tasks. These pragmatic considerations . In particular, using occluded objects to create perspective asymmetries may have unexpected side effects under pragmatic reasoning. Cooperative speakers ought to flexibly adjust their utterances to be more informative uncertainty about what is in the listener's field of view. In a direct replication, we show that scripted confederates violate these expectations by using less informative labels than unscripted directors naturally use. In a follow-up experiment, we directly manipulate whether objects are hidden to test the pragmatic account of speaker informativity in uncertain environments. Hence, apparent failures of visual perspective-taking may in fact be successes for sophisticated theory of mind use in context. Language is well-tuned to deal with knowledge mismatch, and perspective neglect is just part of that.
%Recent theories of language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication. To reconcile these points of view, we use a computational model of communication under uncertainty to analyze the pragmatic demands of a common theory of mind test.  Through a minimal pair of replications comparing scripted vs. unscripted speakers, we show that scripted confederates used in previous studies are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation, thus violating the listener's rational social expectations. This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind. 
\end{abstract}

@@

Our ability to successfully communicate and coordinate with social partners relies on some degree of contextual flexibility in production and interpretation. When trying to refer to a particular dog at a busy dog park, it may be technically correct to call it a ``dog,'' but speakers tend to choose more specific or \emph{informative} expressions, like ``the little terrier with the blue collar'' \citep{VanDeemter16_ComputationalModelsOfReferring,GrafEtAl16_BasicLevel,BrennanClark96_ConceptualPactsConversation,DaleReiter95_GriceanRefs}. Similarly, a hyperbolic statement like ``I waited a million years for a table'' tends to be interpreted as an expression of frustration at longer-than-expected wait times rather than a literal statement about the time elapsed \citep{KaoWuBergenGoodman14_NonliteralNumberWords}.  %A doctor who uses a single Latinate term to tell their colleague about a condition may use many more words to tell a patient the same thing.  % Our success as a cooperative species depends upon our ability to make sense of other agents. 
This systematic contextual variation could in principle be accommodated by internally caching a complex set of different rules. 
But a key insight from the study of pragmatics is that many such phenomena can be more parsimoniously explained as the consequence of shared social reasoning mechanisms \citep{Grice75_LogicConversation, GoodmanFrank16_RSATiCS}. 
Just as making sense of an agent's physical movements in the world requires an accurate model of how beliefs and intentions translate into motor plans \citep{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}, making sense of language depends on an accurate mental model of what a speaker would say, or what a listener would understand, in different situations. 
In other words, communication depends on a flexible \emph{theory of mind} mechanism.

%While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, 
Past efforts to examine theory of mind use in communication have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's knowledge or visual access \citep{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}. 
We argue in this paper that the mental models supporting theory of mind use in communication are richer than previously explored. 
In particular, we consider a class of mental models representing not only a partner's knowledge or visual access but also Gricean expectations of parsimony and cooperativity \citep{FrankeJager16_ProbabilisticPragmatics, Grice75_LogicConversation,Clark96_UsingLanguage,GoodmanFrank16_RSATiCS, FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context. 
%For example, suppose a friend wants to refer to a nearby Dalmatian. 
%There are numerous descriptions that would all be literally true, but different expressions are preferred in different contexts.
%If there are no other dogs nearby, we may expect them to use a parsimonious basic-level term like ``dog.''
%When there is potential for confusion, however, we expect them to consider our likely interpretation and introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \citep{BrennanClark96_ConceptualPactsConversation, VanDeemter16_ComputationalModelsOfReferring, GrafEtAl16_BasicLevel}. 
%When speakers violate these expectations, listeners draw inferences about the intended meaning in order to account for the violation: for example, XXX \cite{XXX, BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}.

@@

This suggests that we consider whether the scripted utterances produced by  confederates in perspective-taking tasks were actually what a speaker in that context would be expected to say \citep{KuhlenBrennan13_LanguageInDialogue}. 
If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
More precisely, we hypothesize that when both players know that objects are occluded in context, pragmatic speakers may add additional precision to their references in order to avoid confusion. 
If a listener reasoning in this way justifiably expects the speaker to be more informative in context, they may \emph{rationally} adopt a ``perspective-neglecting'' strategy. % of choosing the \emph{a priori} more likely referent of the referring expression regardless of occlusion. 
We propose that it is precisely \emph{because} the listener uses theory of mind to form appropriate pragmatic expectations about the speaker's language use that they can be tricked by an uncooperative confederate into misunderstanding. 
%\begin{table*}
%\begin{center}
%\begin{tabular}{ p{2cm} | r | r |  r || r | r | r || r | r | r}
%& \multicolumn{3}{c||}{\% attempted at least once} & \multicolumn{3}{c||}{\% attempted at least twice} & \multicolumn{3}{c}{\% of total cases}\\
%\hline
%& Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2  \\
%Experimental & 71 & 93 & 61 & 46 & 57 & 32 & 30 & 43 & 24\\
%Baseline        & 0   & 7   & 0   & 0   & 0   & 0   & 0   & 2   & 0\\
%\end{tabular}
%\caption{Side-by-side comparison of error rates in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications. The first two sections show the percentage of \emph{participants} attempting to move the occluded distractor at least once, or twice, of the four possible cases. The third section shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object. }
%\label{table:mainResults}
%\end{center}
%\end{table*}
%
%We began by replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment. We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study. We replicated the original finding, but noted a tendency of directors to be overinformative in unscripted filler trials. We then ran the same experiment without using any scripted instructions, observing unconstrained director utterances. We found much greater precision in unconstrained director utterances, which match targets much better than distractors, and better performance of the matchers. This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.

%In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners who are sensitive to the pragmatics of the situation expect these more informative utterances and produce ``errors'' when their expectations are flouted.
In Experiment 1, we conduct a direct replication of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, manipulating whether the speaker is a confederate or naive participant. We find that unscripted speakers naturally produce more informative utterances than those in the confederate's script, and that listeners consequently make fewer errors. In Experiment 2, we manipulate the presence or absence of occlusions in context to directly pinpoint the speaker phenomenon of occlusion sensitivity. We show for the first time that speakers systematically produce longer, more specific utterances than strictly required by their visible context when they have uncertainty about what their partner is seeing. Thus, listeners are rationally justified in ignoring perspective and relying on informative utterances.
Taken together, these results suggest that apparent failures of perspective-taking are in fact attributable to sophisticated pragmatic expectations about speaker behavior---that is, to successful theory of mind use.

 \begin{figure}[t!]
 \begin{center}
 \vspace{-.25cm}
 \includegraphics[scale=.38]{figures/Exp1_design.pdf}
 \vspace{-.5cm}
 \caption{Screenshots of interface on a critical trial: a cassette tape is in view of both players, but a roll of tape is occluded from the speaker's view.}
 \vspace{-.5cm}
 \label{fig:exp1paradigm}
 \end{center}
 \end{figure}

@@

We recruited 350 participants (170 pairs) from Amazon Mechanical Turk to play an interactive communication game using the framework described in \cite{Hawkins15_RealTimeWebExperiments}. Following our preregistered exclusion criteria\footnote{Unless }, we removed k games with 2 or more errors on non-critical items or reported confusion in an exit survey and l games containing non-native english speakers, leaving X games in our final sample.

@@

In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it. Furthermore, even on trials where they moved the correct object, matchers were more likely to hover their cursor over the occluded cell. These results support the claim that listeners neglect the speaker's perspective.

However, when interacting with a natural

@@

%\section{Evaluating models of referring}
%\subsection{Pure egocentrism}
%The simplest strategy for a speaker attempting to refer to a target object $o_i$ is to produce the simplest label that best fits $o_i$, effectively ignoring context and listener knowledge. \todo[inline]{rxdh: technically, mean soft-max here...} Similarly, the simplest strategy for a listener is to pick the object in their view that is best described by the utterance, ignoring the visual access or pragmatic goals of the speaker. These models are purely egocentric, in the sense that they do not engage in any social reasoning. 
%\subsection{Pure consideration of visual access}
%The most immediate model of social reasoning is consideration of the informativeness of a label in a shared context containing many objects. For a speaker, this means preferring labels that are good fits for the target but \emph{not} known distractors in common view. For instance, when referring to a dog in the shared context of other dogs, a speaker may opt to be more informative than simply saying `dog,' since it is ambiguous. For a listener, this means discounting objects that are not in shared context, regardless of how well the description fits. These are not available to the visual access of the speaker and thus not viable referents. 
%\subsection{Probabilistic weighing of perspectives}
%Rather than taking a purely egocentric perspective or purely considering a partner's visual access, speakers or listeners may be probabilistically integrating these two perspectives \cite{HellerParisienStevenson16_ProbabilisticWeighing}. % When conversational expectations lead participants to expect over-informative utterances, they (rationally) place relatively less weight on features of the environment determining common ground, such as shared perceptual access. 
%\subsection{Gricean communication under uncertainty}
%We begin by considering the problem faced by a speaker who wants their partner to select a target object from a context containing other potentially similar objects. %  a simple probabilistic model of reference in the rational speech act (RSA) model. 
%where speakers produce informative utterances by reasoning about a rational listener agent who updates their beliefs about the world according to the literal semantics of the language. This allows us to test pragmatic accounts of task performance. 
%Pragmatic critiques of the director-matcher task point to an additional source of uncertainty that has not previously been explored: uncertainty over a communication partner's visual context. Considering this uncertainty, of course, \emph{is} theory of mind in action; our goal is to show how such a model built on social reasoning can produce the `failures' found by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} as well as the phenomena observed by both \cite{HawkinsGoodman16_Keysar} and \cite{RubioFernandez16_DirectorTaskAttention}. For simplicity, we consider the full $2\times 2$ grid containing a single occluded cell, which was used by \cite{RubioFernandez16_DirectorTaskAttention}. We incorporate this source of uncertainty into both the speaker and listener models as follows:
%\item A pragmatic speaker perceives an array of three objects $\mathcal{O}_S = \{o_i\}$, one of which is identified as a target $t$, but assumes that their partner sees an additional hidden object $o_h$ that is not in common ground. The speaker has uncertainty over the identity of this object, represented by a prior distribution $P(o_h)$ which they marginalize over when deciding what utterance to produce:
%$$S_1(u | t, \mathcal{O}_p) \propto \sum_{o_h} L_0(t| \mathcal{O}_S \cup \{o_h\})P(o_h)P(u)$$
%%\item A pragmatic listener, reasoning about this uncertain speaker, is presented with a full array of four objects $\mathcal{O}_L$, one of which is identified as hidden to the speaker, $o_h$. The listener, however, also has some uncertainty over whether the experimenter is deceptive, represented by a Bernoulli random variable $d$, such that they assign a small prior probability $P(d)$ to the possibility that $o_h$ is actually in common ground. Upon hearing an utterance$u$, they update both their beliefs about which object is the target and whether the experimenter is being deceptive:
%%$$L_2(t, d | u, \mathcal{O}_L) = P(t)\left(S(u | t, \mathcal{O}_L)P(d)\right)^d(S(u | t, \mathcal{O}_L - o_h)(1-P(d))^{1-d}$$
%Technical details are omitted here for brevity, but the full model is fully specified in WebPPL and available to run at %\url{http://forestdb.org/models/keysar.html}. 
%We focus on two critical patterns of results produced by the pragmatic listener $L_2$. First, we consider how it responds to the scripted instruction ``fish'' when presented with a context containing two fish (a red one that is occluded and a blue one in common ground) as well as two other unrelated objects. While it is most likely to choose the red fish in common ground, it has a relatively high probability (43\%) of choosing the hidden red fish, an effect \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} claimed was due to failure of theory of mind. This decision emerges in the model for two primary reasons. For one, a speaker with uncertainty over context could in principle be communicating about any object, and if they believed the hidden object was a red fish, the utterance `fish' would have a high probability of being produced. To a lesser extent, context uncertainty makes the overinformative utterance `blue fish' more likely than `fish', hence the omission of the color modifier has the (subtle) implication of \emph{not} the blue fish. This speaker phenomenon, while subtle under the parameter regime of the current implementation, produces the dynamic observed in the unscripted replication by  \cite{HawkinsGoodman16_Keysar}.
%Second, we consider the listener's posterior over the \emph{deception} Bernoulli variable $d$. Just as \cite{RubioFernandez16_DirectorTaskAttention} found that listeners were more likely to suspect deception after hearing an `overinformative' utterance like `blue fish' when only one fish was in common ground, so too does our pragmatic listener. The pragmatic speaker is relatively more likely to produce the additional color modifier if they see multiple fish than when they see a single fish and simply have uncertainty over the unknown object. Hence, our model simultaneously captures all three phenomena.
